Predicting Solid State Material Platforms for Quantum Technologies

### 2022-06-10 response-letter/response-letter.tex

\input{structure.tex} % Include the file that specifies the document structure

%\longindentation=0pt % Un-commenting this line will push the closing "Sincerely," and date to the left of the page

%----------------------------------------------------------------------------------------
%	YOUR INFORMATION
%----------------------------------------------------------------------------------------

\Who{} % Your name

\Title{  } % Your title, leave blank for no title

\authordetails{%
	Morten Hjorth-Jensen \\ 
	Department of Physics\\ % Your department/institution
	P.O. Box 1048, Blindern\\ % Your address
	N-0316 Oslo, Norway\\ % Your city, zip code, country, etc
	Email: \\ 
	morten.hjorth-jensen \\ 
	@fys.uio.no % Your email address
}

@@

\begin{letter}{
	Dr. Long-Qing Chen\\
	Editor-in-Chief\\
	\textit{npj Computational Materials} \\ 
	npjcompumats@nature.com \\ 
}

%----------------------------------------------------------------------------------------
%	LETTER CONTENT
%----------------------------------------------------------------------------------------

\opening{}

\begin{center}
   \textbf{Resubmission of the manuscript entitled “Predicting Solid State Material Platforms for Quantum Technologies” by O. L. Hebnes, et al.}
\end{center}

@@

\noindent
Dear Editor,

Please find enclosed a revised version of the manuscript entitled “Predicting Solid State Material Platforms for Quantum Technologies” by Oliver L. Hebnes, et al., submitted for publication as an original article in \textit{npj Computational Materials}. There are three files in all: a marked-up Manuscript file, a clean Manuscript file and a Supplementary Information file. Changes are marked in red in the marked-up Manuscript and the Supplementary Information.

We appreciate the opportunity to revise our manuscript. We would also like to thank the reviewers for their careful reading of our manuscript. Their comments point to several places where the manuscript was unclear and have enabled us to substantially improve the paper.

The manuscript has undergone a major revision. Most notable is that we have moved the discussion on the results for the Extended Ferrenti approach from the main text to the Supplementary Information. This was done to diminish the emphasis on this data mining approach. The Ferrenti approach has been kept for several reasons outlined below, but the reasoning behind utilizing this approach has been emphasized, and its benefits clarified. Overall, the text has been shortened wherever possible to address the comment that it was too long, leaving the main text almost a full two pages shorter. All changes to the manuscript are highlighted in red in the marked-up version.

Please find detailed responses to the comments of the Reviewers and Editor below.

%\closing{Sincerely,}
Sincerely,

\noindent Prof. Morten Hjorth-Jensen and Dr. Marianne E. Bathen \\
On behalf of all the authors

@@

The main revision I would like to see addressed is the first point of the second reviewer: What is the utility of the machine learning model if it returns the original classification? The other points they bring up around methodological clarity are important, but providing further justification around the need for the use of machine learning would help reinforce the principle of your study.

\textit{Response: }

\textit{We appreciate the consideration of our manuscript and have carefully addressed all reviewer concerns. The manuscript has undergone a major revision to facilitate the suggested changes. We hope that the methodological clarity has now been improved, and that the logic of the implemented approaches is clear. }

\textit{We  believe that including the Ferrenti approach is important for several reasons.  Firstly, by starting with an established data mining approach we can partly verify the reliability of our methodology before we introduce the empirical approach. The difference in candidate materials extracted from the Ferrenti and empirical approaches also gives insight into the selection criteria and parameters important for suitable qubit materials. Thus, the contrast between the two approaches is important. However, we do realize that the Ferrenti and extended Ferrenti approaches combined put too much emphasis on this data mining approach. A major change to the manuscript that was implemented in order to highlight the empirical approach as the main finding was therefore to move most of the discussion on the extended Ferrenti approach to the Supplementary information. 
The extended Ferrenti approach was initially included because we wanted to remove the effect of practical considerations from the Ferrenti approach, however, this had few detectable effects on the predictions. The same features were recognized as important in the Ferrenti and extended Ferrenti approaches. 
Therefore, the presence of both in the discussion shadows the importance of the empirical approach.
Hence, we have kept mainly the Ferrenti and empirical approaches as discussion points in the main text. Please see the following discussion of this in the main text on page 6, in the Data mining section:} \\ 
\mrk{The findings from both the data mining and machine learning procedures for the extended Ferrenti approach did not differ substantially from those obtained using the Ferrenti approach. This is attributed to the similarities in the selection processes. Therefore, detailed discussion on the findings from the extended Ferrenti approach can be found in the Supplementary Information at [39]. Qualitative conclusions drawn for the Ferrenti approach herein also hold for the extended Ferrenti approach, indicating that the removal of so-called practical considerations did not have a
significant impact on the results. Nonetheless, predictions from the extended Ferrenti approach will be employed below to contrast with and filter the findings from the empirical approach.}

\textit{Secondly, the data mining approach reported by Ferrenti and co-workers follows established expectations (e.g., as reported by Weber et al.) for quantum friendly materials, but no machine learning (ML) has been applied to data sets extracted by this approach. Since ML algorithms are useful for reducing the dimensionality of a data set, it is interesting to see if the ML algorithms recognize the original criteria as the key parameters for labeling the materials. However, we find discrepancies between the selection criteria and the predicted materials (e.g., prediction of NaCl and low band gap materials as suitable).
As such, the expected properties are not able to capture the full physics of the problem. This is in fact an important finding of our work; that the the problem of QT compatibility is more complex than being related to band gap and bonding character alone, and the contrast between the Ferrenti and empirical approaches was necessary to come to this conclusion. However, we realize that this was not sufficiently clear in the original manuscript. 
The following phrase has been added to the Discussion on page 15:} \\
\mrk{As such, the expected properties are not able to capture the full physics of the problem. The contrast between the Ferrenti and empirical approaches reveals that the problem of QT compatibility is more complex than being related to band gap and bonding character alone. Based on our findings we propose that the manifestation of quantum effects in semiconductors is related to the crystal structure symmetry and bonding. }

\textit{Thirdly, the Ferrenti approach can be employed as a filter on the empirical approach. If we consider the materials predicted by both the empirical and Ferrenti approaches as the most suitable, the prediction space is restricted even further. The following paragraph was added to the Discussion section on page 15 to better explain our reasoning:} \\ 
\mrk{It should be noted that performing machine learning on a dataset derived using preconceived notions for which material properties are important may reproduce several of the initial selection criteria. Nonetheless, the Ferrenti approach was included in the present work to highlight expectations from the literature and contrast them with the findings from the empirical one. Additionally, ML methods are often capable of recognizing other patterns than those intended for the data, opening up the possibility that also the Ferrenti approach could yield new insights. Finally, the Ferrenti approaches are employed as a filter on the empirical approach, to provide a better tuned list of candidates for future experimental studies.}

\pagebreak
\noindent
\textbf{Reviewer \#1: }

@@

\textit{Response:}

%\textit{Here ... }

\textit{The Reviewer raises an interesting discussion. We agree that there are portions of our work that should be clarified. We conceive two main themes in the Reviewer's question: (i) the logic of training and testing on the same data set, and (ii) the logic of using preset criteria to design a data set.}

\textit{Firstly, we would like to clarify that the ML methods were not trained and applied to exactly the same data. The criteria that were specified in the Ferrenti approaches left many materials without a label altogether. 
In fact, several materials could not be labeled in the Ferrenti approaches because they did not fall into either of the two categories suitable and unsuitable. 
The labeled data was divided into training and test sets, on which the training and validation of the ML algorithms were performed, and the machine learning to predict new candidate hosts was performed on unlabeled data. Therefore, the training, testing and predicting steps were all performed on different data. This process was repeated for all three approaches, and therefore they exhibit independent training and test data sets.} 
\textit{To clarify this point further we have added the following phrasing to the Data Mining section on page 4:} \\ 
\mrk{The ML methods are trained using the training set and then evaluated on the test
set. Finally, the ML methods are applied to the unlabeled data for which predictions of QT suitability are made.} \\ 
\textit{The notion is repeated on page 10 in the Predicting Suitable Material Hosts for Quantum Technology section:} \\ 
\mrk{After training and validating the ML algorithms on the labeled datasets, the ML methods were applied to unlabeled
data to obtain predictions for suitable QT host materials.}

\textit{Regarding the second point, a principal component analysis does indeed obscure the relationship between the trained machine learning models and the classification criteria. This is one of the reasons why we search for the optimal number of principal components, to observe if any components reveal more information regarding the features. Since all of the parameters from the initial classification criteria are present in the data, including several thousand features generated from Matminer's featurizers, we would still expect the band gap criterion to be absolute, and thus that it  would be reflected in both the optimal principal component model and the machine learning models. However, this is not the case, as the ML-models are achieving an almost perfect accuracy score during the cross-validation, which strengthens our belief  that there are other dimensions in the data that could explain this behavior.} 
\textit{Indeed, as also mentioned by the reviewer, the original Ferrenti criteria are not exactly reproduced by the ML methods. For the Ferrenti and extended Ferrenti approaches, some of the properties we used to guide the labeling are recognized as important by the ML models (e.g., the band gap), but the selection criteria are not fully reproduced. For example, many materials are suggested with band gaps below 0.5~eV in the Extended Ferrenti approach, whereas the selection criterion was set to 1.5~eV. From a materials science perspective, it seems that the exact value of the band gap is not predominant, although the feature itself is important. Furthermore, bases on the criteria set by Weber et al. and Ferrenti et al. we expected covalent materials to dominate, but NaCl is predicted as suitable in the Ferrenti approach. The ML models will naturally weight the criteria differently from the specifications as they are not aware of the criteria, and seem to be recognizing some pattern in the data that was not exactly anticipated. We have attempted to emphasize this point in the revised version of the manuscript as also mentioned above, please see the response to the Editor's comment. The topic is discussed  in the following paragraph on page 11 in the Predicting Suitable Material Hosts for Quantum Technology  section:} \\ 
\mrk{The ML methods predict materials as suitable that are not expected according to, e.g., Ref. [8]. Indeed, NaCl is predicted as a suitable candidate to minimum confidences of 0.83 and 0.61 for two different configurations, despite the strong electrostatic interactions between Na and Cl and the ionic character of their bonding.}  \\ 
\textit{and in the Discussion section on page 13:} \\ 
\mrk{Taking a closer look at the reasoning behind the choices made by the different ML methods during the classification process, we can start to identify important driving forces for manifestation of quantum compatible properties in
semiconductors. The analysis of the principal components extracted from the ML methods revealed that the most important principal component for the Ferrenti approach encompasses features related to the band gap and chemical environment. This means that the band gap criterion imposed in the training set selection is at least somewhat satisfied. The Ferrenti approach does not entirely reproduce the logic of the initial selection process, however, as several low band gap (< 0.5 eV) materials were highlighted as suitable by the ML methods. For the empirical approach, on the other hand, band gap related features were not recognized as important in the dominant principal component.}

\textit{Thirdly, as mentioned above, a driving force behind selecting to report on the Ferrenti approaches was the comparison to the empirical one. The selection criteria for the Ferrenti approach were based on the general  expectations from the field. Hence, an analysis from the ML perspective on which features this would entail (i.e., from Matminer) was necessary to compare with the findings from the empirical approach, which consistute the main findings of this work. Finally, the agreement in predictions from the three approaches could lend weight to a prediction and merit further experimental studies. In other words, the Ferrenti approaches are used as a filter for the empirical one. }

\textit{Importantly, the above considerations were not sufficiently clear in the original version of the manuscript. Therefore, we have tried to make these three points more clear in the overall text. Furthermore, the removal of the extended Ferrenti approach from the main text should hopefully also clarify the juxtaposition of the Ferrenti and empirical approaches. Please also see the response to the Editor above and the accompanying changes to the manuscript.}

@@

Reviewer \#2: \\
In the machine-learning models trained on the extended Ferrenti approach, the authors note that the models predict unsuitable materials to be suitable, e.g. NaCl. How were these materials categorized by the original extended Ferrenti criteria? It is not necessarily clear that the Ferrenti/extended Ferrenti criteria listed on pages 5 and 6 of this work would reliably categorize ionic materials with cubic symmetry as unsuitable, so it is unsurprising that machine-learning models trained on such data would have these problems.

\textit{Response:} 
\textit{The material NaCl involves two elements where more than half does not have a natural abundance of spin zero isotopes, and is therefore excluded from being labeled as suitable. In addition, NaCl is excluded from being labeled as unsuitable from the criteria due to its nonmagnetic label in Materials Project. We agree that it is not necessarily clear that either the Ferrenti or the extended Ferrenti approaches would categorize ionic materials with cubic symmetry as unsuitable, but it is clear from these criteria why NaCl is excluded from the labeling process. Therefore, NaCl is found unlabeled in the test data, while the ML algorithms in both Ferrenti and extended Ferrenti altogether predicts the material as suitable.}  
\textit{The following note was added to this effect on page 11 in the Predicting Suitable Material Hosts for Quantum Technology section:} \\ 
\mrk{Note that NaCl was excluded from being labeled as both unsuitable and suitable in the Ferrenti approach and was therefore found in the unlabeled dataset. For the empirical approach, NaCl was included in the labeled data in the training set as an unsuitable candidate. }

@@

Reviewer \#2: \\
In the Ferrenti and extended Ferrenti searches, the authors also state that they only include materials that have been calculated to be “non-magnetic”. However, it is unclear how this criterion would be implemented in practice: while ferromagnetic materials are usually easy enough to identify, other types of magnetic order such as antiferromagnetism can be more difficult. Many of the automated calculations inside these databases are based on the primitive cell, which might not be large enough to be able to properly represent the antiferromagnetic configuration. Therefore, the authors should clearly define their criteria for determining if a material is non-magnetic, e.g. including checks to see if larger cells have been explored.

\textit{Response:}

\textit{The reviewer raises an important issue. 
The data regarding a material's magnetic character is extracted from the Materials Projects database. Indeed, the majority of these calculations are based on the primitive cell, however, Materials Project performs an initial relaxation of cell and lattice parameters using a $1000$ / (number of atoms in the cell) k-point mesh to ensure that all properties calculated are representative of the idealized unit cell for each respective structure. As a result, we can find Fe (https://materialsproject.org/materials/mp-13/) labeled as ferromagnetic and NiO (https://materialsproject.org/materials/mp-19009/) as antiferromagnetic in our data set.}

\textit{Materials Project also includes larger structures of the same material. We have included all structures and materials from Materials Project, but we have not checked which materials are represented as a larger cell in our data. We have thereby not verified whether antiferromagnetic ordering has been investigated for all cases. This is an improvement that could be made to our method in future studies.}

\textit{These points were not sufficiently clear in the original manuscript which has been remedied in the revised version. The following sentence was added to the Data Mining section on page 5 to clarify the part about ferromagnetic ordering: } \\
\mrk{Note that larger cells are sometimes needed to verify antiferromagnetic ordering so the criteria mainly target ferromagnetic ordering under the labels magnetic/non-magnetic.}

\textit{The following clarifications were added to the Methods section under Databases:} \\ 
\mrk{The data regarding a material's magnetic character is extracted from the Materials Projects database. Indeed, the majority of these calculations are based on the primitive cell, however, Materials Project performs an initial relaxation of cell and lattice parameters using a $1000$ / (number of atoms in the cell) k-point mesh to ensure that all properties calculated are representative of the idealized unit cell for each respective structure. As a result, we can find, e.g., Fe labeled as ferromagnetic and NiO as antiferromagnetic in our dataset. Furthermore, MP contains structures of varying sizes for the same material. We have included all structures and materials from the Materials Project, but we have not checked which materials are represented as a larger cell in our data. We have thereby not verified whether antiferromagnetic ordering has been investigated for all cases. This is an improvement that could be made to our method in a future study.}

@@

Reviewer \#2: \\
In the empirical approach, the data set is constructed based on the materials that are currently used in quantum technologies. The authors should discuss if there is any intrinsic bias in this data set, e.g. that these materials are selected based on what are readily available to researchers in this field, rather than the materials that are inherently the most suitable.

\textit{We agree with the Reviewer. Availability of affordable and high quality crystals has of course limited the search for quantum compatible defects and solid state effects in semiconductors. Moreover, it is likely that the discovery of single photon emission and spin manipulation for the NV center in diamond has also guided the search for similar properties in other materials, with the search commencing with similar materials to diamond, such as different polytypes of SiC. The criteria of Weber et al. (Ref.~[8] of the main text) are also interesting to consider in this regard, in particular the assumption that wide band gap materials are preferable. The recent discovery of telecom compatible SPEs in Si (Refs.~[54-55] in the main text), a lower band gap material, show that this may have been a premature assumption. We hope that our work can contribute to widening the search for quantum compatible semiconductor platforms. 
A discussion on this has now been added to the Data Mining section on page 7: } \\ 
\mrk{Note that there is some potential for inherent bias in the dataset, in part due to experimental work being limited by the availability and cost of materials and processing. Moreover, the discovery of the quantum compatible properties of the NV center in diamond naturally led early searchers to comparable materials such as silicon carbide. }

@@

Reviewer \#2: \\
The authors use the Materials Project as their primary database to search for ICSD entries. They then search for similar entries in other databases (OQMD, JARVIS-DFT, Citrination, etc.), and combine the data. However, it is unclear what additional data they are extracting from the other databases. How do they identify/define similar materials – are they based on the same ICSD number, on comparing the structures, etc.? If two different databases have different values for the same property, e.g. two different values of the band gap (perhaps due to different U parameters being used), how do they decide which value to use?

\textit{Response:}

\textit{From the initial criterion, we define that a material is required to have an ICSD identifier (ID) in Materials Project. This ID is included in the AFLOW, AFLOW-ML, JARVIS-DFT and OQMD databases as well. Due to the various different techniques of approximating the bandgap, either demonstrated experimentally or estimated using the DFT-calculations with the functionals PBE or PBE+U, and as we do not exactly know which method is most accurate, we do not choose one but rather let them all be represented as columns (features) in our data. 
%We believe information from other databases should be handled as supplements rather than choosing one. 
We note that other approaches to this issue  exist, such as letting the same material be represented as two entries (rows) in our data but with different values for the same property. However, such replications could substantially increase the size and complexity of the current experiment, while not necessarily resulting in any credible benefits.}

@@

%If two different databases have different values for the same property, we employed the value from ... }

### 2022-06-10 revision/revised-paper_marked-up.tex
Updates from Overleaf

\newcommand{\mrk}[1]{\textcolor{red}{#1}}

@@

Semiconductor materials provide a compelling platform for quantum technologies (QT), and the properties of a vast amount of materials can be found in databases containing information from both experimental and theoretical explorations. However, searching these databases to find promising candidate materials for quantum technology applications 
is a major challenge. Therefore, we have developed a framework for the automated discovery of semiconductor host platforms for QT using material informatics and machine learning methods, resulting in a dataset consisting of over \mrk{$25,000$} materials and nearly $5000$ physics-informed features. Three approaches were devised, named the Ferrenti, extended Ferrenti and the empirical approach,  to label data for the supervised machine learning (ML) methods logistic regression, decision trees, random forests and gradient boosting. 
We find that of the three, the empirical approach relying exclusively on findings from the literature predicted substantially fewer candidates than the other two approaches with a clear distinction between suitable and unsuitable candidates when comparing the two largest eigenvalues in the covariance matrix. 
In contrast to expectations from the literature and that found for the Ferrenti and extended Ferrenti approaches focusing on band gap and ionic character, the ML methods from the empirical approach highlighted features related to symmetry and crystal structure, including bond length, orientation and radial distribution, as influential when predicting a material as suitable for QT. 
All three approaches and all four ML methods agreed on a subset of $47$ eligible candidates 
%(to a probability of $>50 \ \%$) 
of $8$ elemental, $29$ binary, and $10$ tertiary compounds, and provide a basis for further material explorations towards quantum technology.

@@

\section*{Introduction}
Quantum technologies (QT) based on solid state platforms have attracted a lot of attention during \mrk{recent}  years. 
Among the promising applications that are already available we find important breakthroughs such as in vivo sensing of magnetic fields in cells \cite{Lesage_2013}, secure communication over large distances by separation of entangled photons \cite{Ursin2007} and, notably, various quantum information processing prototypes and architectures \cite{Arute_2019}.  
Quantum computers are in high demand to meet the increasing need for computing power to solve complex and high-dimensional scientific problems. 
%that are beyond the capability of existing high-performance computing %facilities.   
Recent advances have highlighted the potential of quantum information processors to outperform state-of-the art high-performance computing facilities. Indeed, Google's $53$ qubit quantum computer based on superconducting electronics solved a computational problem that was beyond the capabilities of a \mrk{$200,000$} core supercomputer \cite{Arute_2019}. Most recently, IBM announced its $127$ qubit quantum processor~\cite{IBM2021}. Simultaneously, the concepts of entanglement and teleportation may eventually facilitate advanced quantum communication protocols such as quantum cryptography and the quantum internet \cite{quantum-internet-kimble}, spurring further investigations into technologies based on quantum mechanics.

@@

Here we provide a framework for the data mining and automated discovery of promising \mrk{semiconductor hosts for QT} using targeted database search and ML methods combined with knowledge from the field. 
Analyzing the output of the ML methods reveals that, given a suitable initial set of labeled materials for training and testing, it is possible to discern the physical mechanisms that govern a material's suitability for quantum applications.  
In this framework we do not distinguish between the specific mechanism giving rise to properties such as single-photon emission and long spin coherence times (e.g., semiconductor defects or quantum dots); instead, we attempt to target all materials that may accommodate the desired characteristics.  
The methodology developed herein can be modified for other material types and application areas provided that high quality databases containing relevant theoretical and/or experimental data are available.

The developed procedure relies on data extraction from the databases Materials Project \cite{Jain2013,Jain2018}, the Open Quantum Materials Database (OQMD) \cite{Saal2013, Kirklin2015}, JARVIS-DFT \cite{Choudhary2020}, AFLOW \cite{Curtarolo2012, Curtarolo2012a, Calderon2015} and AFLOW-ML \cite{Isayev2017}. 
The Matminer Python library for data mining \cite{Ward2018} was then used for material analysis to featurize the extracted data. An important aspect of \mrk{the} work is the database building and pertinent development of the datasets for the ML methods. \mrk{Three different approaches to data mining were devised}: (i) \emph{the Ferrenti approach} which is similar to that proposed by \citeauthor{Ferrenti2020} \cite{Ferrenti2020}, (ii) \emph{the extended Ferrenti approach} and (iii) \emph{the empirical approach}. \mrk{The} two first methodologies base their data extraction protocols on broad material descriptors, leading to large sets of potentially suitable candidates \cite{Mehta2019,Hastie2009}. The empirical approach, on the other hand, relies on including materials with experimentally proven advantageous characteristics in the training set and therefore yields a narrower set of possible candidates. The three resulting sets of labeled data were then analyzed with the four supervised ML methods logistic regression, decision trees, random forests and gradient boosting \cite{Hastie2009,Murphy2012}, yielding \mrk{$47$} predicted candidates that were common between all approaches and ML methods.  
Example materials \mrk{among the predictions} include ZnGeP$_2$, CdS, BP, BC$_2$N, GeC and InP. Focused theoretical and experimental studies, using methods such as, e.g., density functional theory, photoluminescence spectroscopy and optically detected magnetic resonance, are needed to verify the predictions that quantum properties may manifest as a result of defects or nanostructures in the above listed materials. 
%Write better! 
Importantly, our findings also reveal which material properties are weighted by the ML methods upon predicting a material as suitable for QT applications, thereby opening up for new  discoveries in the field of quantum technologies.

@@

\subsection*{Information Flow}

The information stream in this work can be regarded as many inter-connected modular parts. 
The initial step for gathering material data and building features is visualized by the outer flowchart in Fig.~\ref{fig:flowchart} (Jupyter notebooks containing the full 
workflow can be found at \cite{Ohebbi2021}).
We start by extracting all entries in the Materials Project (MP) database  \cite{Jain2013,Jain2018} that match a specific query. 
The MP database contains ground state properties of different materials that are computed using density functional theory. The DFT calculations in the database were performed using the Vienna {\em ab initio} simulation package (VASP) \cite{Kresse1996} and the Perdew-Burke-Ernzerhof (PBE) \cite{Perdew1996} exchange-correlation functional to calculate the electronic structure of the materials. 
We note that despite being immensely successful in describing a number of material properties, PBE is widely known for underestimating the band gap of semiconductors \cite{Freysoldt2014}. Therefore, not all properties predicted using the PBE functional are reliable, and the band gaps in particular cannot be trusted in absolute terms. Nonetheless, the functional is in wide use due to the combination of reasonable accuracy and high computational throughput, and is usually considered to be reliable for large-scale trends in semiconductor \mrk{material} properties.

\begin{figure}[t]
    \centering
    \includegraphics[width=0.9\textwidth]{figures/flow_chart_2.png}
    \caption{The project workflow starting from an initial Materials Project (MP) query, and ending with a featurized dataset with entries from several other databases. 
    To limit the memory and computational usage, the data are partitioned into smaller subsets where the respective Pymatgen objects (Pymatgen is a robust, open-source Python library for materials analysis \cite{pymatgen}) are obtained through a query to be used in the following featurization steps. This process is repeated iteratively until all the data has been featurized. \emph{DOS} refers to density of states and \emph{band} to the electronic band structure. 
    }
    \label{fig:flowchart}
\end{figure}

The conditions for the initial MP query are that the materials must derive from the Inorganic Crystal Structure Database (ICSD) and have a band gap wider than $0.1$~eV to exclude metallic compounds. The ICSD is the world's largest database for completely identified inorganic crystal structures  \cite{Allen1987,Zagorac2019}. In a parallel step, entries that are deemed similar to the entries from the initial query are extracted from the databases OQMD  \cite{Saal2013,Kirklin2015}, JARVIS-DFT \cite{Choudhary2020}, AFLOW \cite{Curtarolo2012, Curtarolo2012a, Calderon2015}, AFLOW-ML \cite{Isayev2017} and the Citrination platform \cite{OMaraJordan2016MDIA}. The results of these steps are combined into a dataset for further analysis.

After material extraction we apply tools from the open-source library Matminer \cite{Ward2018} to generate thousands of features from the data. We will refer to this process as featurization. A schematic visualization of the featurization process in Matminer is shown in  Fig.~\ref{fig:flowchart} 
and focuses on a material's composition,  structure, atomic sites, density of states and band structure. 
The $39$ featurizers (each generates several features) selected as material descriptors in this work are \mrk{described} in the Supplementary Information at \cite{supplementary}. The selection of features was kept rather wide \mrk{to avoid making} {\em ex ante} assumptions on which features best describe a solid state material platform for \mrk{QT}.

The constructed dataset encompasses compounds formed by a plethora of combinations of surfaces, interfaces, nanostructures, compositions and structures, but this complexity is not necessarily reflected in the material descriptors. 
Furthermore, we have utilized data obtained from high-throughput density functional theory calculations. Indeed, there are possible errors associated with every step, starting from an initial calculation, adding of data in the database, gathering of data, featurization of data, preprocessing of data, data mining, and finally training a model and making a prediction. Unfortunately, if an error has \mrk{occurred} in the first part of the process, the error \mrk{is carried along} and will get increasingly harder to detect. Thus, \mrk{the} results depend on the quality of data in the \mrk{employed} databases.

>>>>

\newcommand{\mrk}[1]{\textcolor{red}{#1}}
\usepackage[useregional]{datetime2}

@@

Semiconductor materials provide a compelling platform for quantum technologies (QT), and the properties of a vast amount of materials can be found in databases containing information from both experimental and theoretical explorations.
However, searching these databases to find promising candidate materials for quantum technology applications is a major challenge. 
Therefore, we have developed a framework for the automated discovery of semiconductor host platforms for QT using material informatics and machine learning methods, resulting in a dataset consisting of over \mrk{\num{25000}} materials and nearly \num{5000} physics-informed features.
\mrk{Three different approaches were devised} to label data for the supervised machine learning (ML) methods logistic regression, decision trees, random forests and gradient boosting. 
We find that \mrk{among these, the} empirical approach relying exclusively on findings from the literature \mrk{exhibited the clearest  distinction}  between \mrk{predicted} suitable and unsuitable candidates when comparing the two largest eigenvalues in the covariance matrix. 
In contrast to expectations from the literature  %and that found for the Ferrenti and extended Ferrenti approaches focusing on band gap and ionic character, 
\mrk{focusing on band gap and ionic character as important features for QT compatibility,}  
the ML methods from the empirical approach highlighted features related to symmetry and crystal structure, including bond length, orientation and radial distribution, as influential when predicting a material as suitable for QT. 
All three approaches and all four ML methods agreed on a subset of $47$ eligible candidates 
%(to a probability of $>50 \ \%$) 
of $8$ elemental, $29$ binary, and $10$ tertiary compounds, and provide a basis for further material explorations towards quantum technology.

@@

\section*{Introduction}
Quantum technologies (QT) based on solid state platforms have attracted a lot of attention during \mrk{recent}  years. 
Among the promising applications that are already available we find important breakthroughs such as in vivo sensing of magnetic fields in cells \cite{Lesage_2013}, secure communication over large distances by separation of entangled photons \cite{Ursin2007} and, notably, various quantum information processing prototypes and architectures \cite{Arute_2019}.  
Quantum computers are in high demand to meet the increasing need for computing power to solve complex and high-dimensional scientific problems. Recent advances have highlighted the potential of quantum information processors to outperform state-of-the art high-performance computing facilities.
Indeed, Google's $53$ qubit quantum computer based on superconducting electronics solved a computational problem that was beyond the capabilities of a \mrk{\num{200000}} core supercomputer \cite{Arute_2019}.
Most recently, IBM announced its $127$ qubit quantum processor~\cite{IBM2021}. Simultaneously, the concepts of entanglement and teleportation may eventually facilitate advanced quantum communication protocols such as quantum cryptography and the quantum internet \cite{quantum-internet-kimble}, spurring further investigations into technologies based on quantum mechanics.

@@

\mrk{Herein, a framework is provided} for the data mining and automated discovery of promising \mrk{semiconductor hosts for QT} using targeted database search and ML methods combined with knowledge from the field. 
Analyzing the output of the ML methods reveals that, given a suitable initial set of labeled materials for training and testing, it is possible to discern the physical mechanisms that govern a material's suitability for quantum applications.  
\mrk{This framework does not} distinguish between the specific mechanism giving rise to properties such as single-photon emission and long spin coherence times (e.g., semiconductor defects or quantum dots); instead, we attempt to target all materials that may accommodate the desired characteristics.  
The methodology developed herein can be modified for other material types and application areas provided that high quality databases containing relevant theoretical and/or experimental data \mrk{is} available.

The developed procedure relies on data extraction \mrk{from different databases and the subsequent featurization of the data.} 
%from the databases Materials Project \cite{Jain2013,Jain2018}, the Open Quantum Materials Database (OQMD) \cite{Saal2013, Kirklin2015}, JARVIS-DFT \cite{Choudhary2020}, AFLOW \cite{Curtarolo2012, Curtarolo2012a, Calderon2015} and AFLOW-ML \cite{Isayev2017}. 
%The Matminer Python library for data mining \cite{Ward2018} was then used for material analysis to featurize the extracted data. 
An important aspect of \mrk{the} work is the database building and pertinent development of the datasets for the ML methods. \mrk{Three different approaches to data mining were devised}: (i) \emph{the Ferrenti approach} which is similar to that proposed by \citeauthor{Ferrenti2020} \cite{Ferrenti2020}, (ii) \emph{the extended Ferrenti approach} and (iii) \emph{the empirical approach}. \mrk{The} two first \mrk{data extraction protocols are based on} broad material descriptors, leading to large sets of potentially suitable candidates \cite{Mehta2019,Hastie2009}. The empirical approach, on the other hand, relies on including materials with experimentally proven advantageous characteristics in the training set and therefore yields a narrower set of possible candidates. The three resulting sets of labeled data were then analyzed with the four supervised ML methods logistic regression, decision trees, random forests and gradient boosting \cite{Hastie2009,Murphy2012}, yielding \mrk{$47$} predicted candidates that were common between all approaches and ML methods.
Example materials \mrk{among the predictions} include ZnGeP$_2$, CdS, BP, BC$_2$N, GeC and InP. Focused theoretical and experimental studies 
%, using methods such as, e.g., density functional theory, photoluminescence spectroscopy and optically detected magnetic resonance, 
\mrk{are needed} to verify the predictions that quantum properties may manifest as a result of defects or nanostructures in the above listed materials. 
%Write better! 
Importantly, our findings also reveal which material properties are weighted by the ML methods upon predicting a material as suitable for QT applications, thereby opening up for new  discoveries in the field of quantum technologies.

@@

\subsection*{Information Flow} 
The information stream in this work can be regarded as many inter-connected modular parts. 
The initial step for gathering material data and building features is visualized by the outer flowchart in Fig.~\ref{fig:flowchart} (Jupyter notebooks containing the full 
workflow can be found at \cite{Ohebbi2021}).
We start by extracting all entries in the Materials Project (MP) database  \cite{Jain2013,Jain2018} that match a specific query.  
The MP database contains ground state properties of different materials that are computed using density functional theory. The DFT calculations in the database were performed using the Vienna {\em ab initio} simulation package (VASP) \cite{Kresse1996} and the Perdew-Burke-Ernzerhof (PBE) \cite{Perdew1996} exchange-correlation functional to calculate the electronic structure of the materials. 
We note that despite being immensely successful in describing a number of material properties, PBE is widely known for underestimating the band gap of semiconductors \cite{Freysoldt2014}. Therefore, not all properties predicted using the PBE functional are reliable, and the band gaps in particular cannot be trusted in absolute terms. Nonetheless, the functional is in wide use due to the combination of reasonable accuracy and high computational throughput, and is usually considered to be reliable for large-scale trends in semiconductor \mrk{material} properties.

\begin{figure}[t]
    \centering
    \includegraphics[width=0.9\textwidth]{figures/flow_chart_2.png}
    \caption{The project workflow starting from an initial Materials Project (MP) query, and ending with a featurized dataset with entries from several other databases. 
    To limit the memory and computational usage, the data \mrk{is} partitioned into smaller subsets where the respective Pymatgen objects (Pymatgen is a robust, open-source Python library for materials analysis \cite{pymatgen}) are obtained through a query to be used in the following featurization steps. This process is repeated iteratively until all the data has been featurized. \emph{DOS} refers to density of states and \emph{band} to the electronic band structure. 
    }
    \label{fig:flowchart}
\end{figure}

The conditions for the initial MP query are that the materials must derive from the Inorganic Crystal Structure Database (ICSD) and have a band gap wider than \SI{0.1}{\electronvolt} to exclude metallic compounds.
The ICSD is the world's largest database for completely identified inorganic crystal structures  \cite{Allen1987,Zagorac2019}. In a parallel step, entries that are deemed similar to the entries from the initial query are extracted from the \mrk{following databases: the Open Quantum Materials Database (OQMD)}  \cite{Saal2013,Kirklin2015}, JARVIS-DFT \cite{Choudhary2020}, AFLOW \cite{Curtarolo2012, Curtarolo2012a, Calderon2015}, AFLOW-ML \cite{Isayev2017} and the Citrination platform \cite{OMaraJordan2016MDIA}. The results of these steps are combined into a dataset for further analysis.

After material extraction \mrk{tools from the open-source library Matminer \cite{Ward2018} were applied} to generate thousands of features from the data. We will refer to this process as featurization. A schematic visualization of the featurization process in Matminer is shown in  Fig.~\ref{fig:flowchart} 
and focuses on a material's composition,  structure, atomic sites, density of states and band structure. 
The $39$ featurizers (each generates several features) selected as material descriptors in this work are \mrk{described} in the Supplementary Information at \cite{supplementary}. The selection of features was kept rather wide \mrk{to avoid making} {\em ex ante} assumptions on which features best describe a solid state material platform for \mrk{QT}.

The constructed dataset encompasses compounds formed by a plethora of combinations of surfaces, interfaces, nanostructures, compositions and structures, but this complexity is not necessarily reflected in the material descriptors. 
Furthermore, \mrk{the extracted data was}  obtained from high-throughput density functional theory calculations. Indeed, there are possible errors associated with every step, starting from an initial calculation, \mrk{adding of data in the database, featurization and preprocessing the data, data mining and labeling, and finally training a model and making a prediction.}  Unfortunately, if an error \mrk{occurrs} in the first part of the process, \mrk{it will be  carried along and get} increasingly harder to detect. Thus, \mrk{the output strongly depends}  on the quality of data in the \mrk{employed} databases.
